Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2020-12-10 GWLDCII.tex
a few initial comments, will hopefully edit more later today

\title{Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators}
\author{Susama Agarwala\thanks{SA was partially supported by an Office of Naval Research grant.}, Si\^an Fryer, and Karen Yeats\thanks{KY is supported by an NSERC Discovery grant, by the Canada Research Chair program, and though some the time this work was developed also by a Humboldt Fellowship from the Alexander von Humboldt foundation.}}
\date{}

@@

In contrast, the first paper in this series focuses on combinatorial and matroidal properties of the Wilson loop diagrams. The main focus of that paper is enumerating the positroid cells that are associated to Wilson loop diagrams with a fixed number of vertices and propagators, as well as counting the number of Wilson loop diagrams that map to the same positroid. The results of the first paper are not a prerequisite for this paper.

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite[section 11]{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrands associated to the BCFW diagrams define a volume form on the Amplituhedron. This gives a novel and elegant interpretation of scattering amplitudes for this theory, as well as a means of greatly simplifying the calculations involved in traditional Feynman diagram approaches. Since then, there has been significant work to understand the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrands associates a volume form to a convex space different from those in the decomposition the Amplituhedron \cite{CachazoSvrcekWitten, Hodges2013}.  \hlfix{The Wilson loop diagrams discussed in this paper are, loosely speaking, dual graphs of the $N^kMHV$ diagrams. While the Wilson loop diagrams contain the exact same information as the $N^kMHV$ diagrams for the purposes of understanding $N=4 SYM$ theory, this dual perspective is better suited to understanding the combinatorics and geometry associated to the theory.}{added to address request (3) of report M} Therefore, we part with the notation in the physics literature, and name these dual graphs of $N^kMHV$ diagtams Wilson loop diagrams in order to emphasize the combinatorial and geometric significane of these object. In \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

In the above work and elsewhere, it is clear that while the space filled by the Wilson loop diagrams is related to the Amplituhedron, they are not the same space. For instance, it is conjectured that the Amplituhedron is orientable, whereas Agarwala and Marcott show in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams as a subspace of $\Gr(k, n+1)$ can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, which can be seen by comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. This is an intriguing observation, as it is conjectured that the space parametrized by the Wilson loop diagrams will project down to something close to the Amplituhedron \cite{wilsonloop, Amplituhedronsquared}.

It is worth noting at this point that the combinatorics of the Wilson loop diagrams are different from the combinatorics of the Amplituhedron, just as the the volume forms associated to the BCFW diagrams are different from those associated to the Wilson Loop diagrams. For instance, the BCFW are most readily related to the geometry of a postive Grassmanian via plabic graphs. Meanwhile, the Wilson loop diagrams correspond to matrices with variable valued entries (see section \ref{section WLD background}) and Grassmann Necklaces (see Algorithm \ref{alg:put GN on WLD}).  These two combinatorial representations of the positive Grassmannians have different strengths and weaknesses, and it is not possible to give a one to one translation of the combinatorial tools used for the Amplituhedron and BCFW diagrams, and the combinatorial tools used for the Wilson loop diagrams, associated matrices and Grassmann Necklaces. For instance, square moves and merge-expand moves give different graphical representations of the same positroid cell. There is no equivalent diagramatical equivalence for Wilson loop diagrams, though it has been shown (see \cite{wilsonloop, GWLDCI}) that certain classes of diagrams parametrize the same positroid cells. On the other hand, in order to get a parametrization of the associated positroid cell from the plabic graphs, one must assign weights to the vertices and used them to compute Pl\"{u}cker coordinates \cite[Section 11]{Postnikov}, while the Wilson loop diagrams \emph{are} a parametrization of the associated positroid. I.e. one can compute the Pl\"{u}cker coordinates directed from the associated matrices.  \sanote{This para is in answer to 1 and 4 of report Y. Should it be moved (in a modified form) to the begining of section 3? Should it be in both places?}

>>>>

\title{Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators}
\author{Susama Agarwala\thanks{SA was partially supported by an Office of Naval Research grant.}, Si\^an Fryer, and Karen Yeats\thanks{KY is supported by an NSERC Discovery grant, by the Canada Research Chair program, and also, through some the time this work was developed, by a Humboldt Fellowship from the Alexander von Humboldt foundation.}}
\date{}

@@

In contrast, the first paper in this series focuses on combinatorial and matroidal properties of the Wilson loop diagrams. The main focus of that paper is on enumerating the positroid cells that are associated to Wilson loop diagrams with a fixed number of vertices and propagators, as well as counting the number of Wilson loop diagrams that map to the same positroid. The results of the first paper are not a prerequisite for this paper.

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite[section 11]{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrands associated to the BCFW diagrams define a volume form on the Amplituhedron. This gives a novel and elegant interpretation of scattering amplitudes for this theory, as well as a means of greatly simplifying the calculations involved in traditional Feynman diagram approaches. Since then, there has been significant work devoted to understanding the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrands associates a volume form to a convex space different from those in the decomposition the Amplituhedron \cite{CachazoSvrcekWitten, Hodges2013}.  \hlfix{The Wilson loop diagrams discussed in this paper are, loosely speaking, dual graphs of the $N^kMHV$ diagrams. While the Wilson loop diagrams contain the exact same information as the $N^kMHV$ diagrams for the purposes of understanding $N=4 SYM$ theory, this dual perspective is better suited to understanding the combinatorics and geometry associated to the theory.}{added to address request (3) of report M -- let's discuss, I don't think this makes sense as written} Therefore, \hlfix{we part with the notation in the physics literature, and name these dual graphs}{I'm not sure that's true} of $N^kMHV$ diagrams Wilson loop diagrams in order to emphasize the combinatorial and geometric significane of these object. In \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in \hlfix{this}{What does this ``this'' refer to, also this sentence should probably start a new paragraph} literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

In the above work and elsewhere, it is clear that while the space \hlfix{filled by}{``filled by'' vs ``parameterized by'' in the previous paragraph} the Wilson loop diagrams is related to the Amplituhedron, they are not the same space. For instance, it is conjectured that the Amplituhedron is orientable, whereas Agarwala and Marcott show in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams as a subspace of $\Gr(k, n+1)$ can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, which can be seen by comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. This is an intriguing observation, as it is conjectured that the space parametrized by the Wilson loop diagrams will project down to something close to the Amplituhedron \cite{wilsonloop, Amplituhedronsquared}.

It is worth noting at this point that the \hlfix{combinatorics of the Wilson loop diagrams are different from the combinatorics of the Amplituhedron}{I think we should discuss this paragraph -- it may be misleading in a bad way}, just as the the volume forms associated to the BCFW diagrams are different from those associated to the Wilson Loop diagrams. For instance, the BCFW are most readily related to the geometry of a postive Grassmanian via plabic graphs. Meanwhile, the Wilson loop diagrams correspond to matrices with variable valued entries (see section \ref{section WLD background}) and Grassmann Necklaces (see Algorithm \ref{alg:put GN on WLD}).  These two combinatorial representations of the positive Grassmannians have different strengths and weaknesses, and it is not possible to give a one to one translation of the combinatorial tools used for the Amplituhedron and BCFW diagrams, and the combinatorial tools used for the Wilson loop diagrams, associated matrices and Grassmann Necklaces. For instance, square moves and merge-expand moves give different graphical representations of the same positroid cell. There is no equivalent diagramatical equivalence for Wilson loop diagrams, though it has been shown (see \cite{wilsonloop, GWLDCI}) that certain classes of diagrams parametrize the same positroid cells. On the other hand, in order to get a parametrization of the associated positroid cell from the plabic graphs, one must assign weights to the vertices and used them to compute Pl\"{u}cker coordinates \cite[Section 11]{Postnikov}, while the Wilson loop diagrams \emph{are} a parametrization of the associated positroid. I.e. one can compute the Pl\"{u}cker coordinates directed from the associated matrices.  \sanote{This para is in answer to 1 and 4 of report Y. Should it be moved (in a modified form) to the begining of section 3? Should it be in both places?}
